Can Auxiliary Indicators Improve COVID-19 Forecasts and Hotspot Prediction?

### 2021-06-15 forecast/paper/paper.tex
Fix logistic typo

We treat hotspot prediction as a binary classification problem and use a setup 
altogether quite similar to the forecasting setup described previously.
Separately for each $a=7,\ldots,28$, to predict $Z_{\ell,t+a}$, we consider a
simple logistic model:
\begin{equation}
\label{eq:hotspot_ar}
\logit(Z_{\ell,t+a} \,|\, Y_{\ell,s}, s \leq t)  
= \alpha^{a,\tau} + \sum_{j=0}^2 \^{a,\tau}_j Y^\Delta_{\ell,t-7j}.  
\end{equation}
In addition to this ``pure'' autoregressive model, we also consider five
logistic models of the form: 
\begin{multline}
\label{eq:hotspot_ar_x}
\logit(Z_{\ell,t+a} \,|\, Y_{\ell,s}, X_{\ell,s}, s \leq t)  
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j} +  
\sum_{j=0}^2 \gamma^{a,\tau}_j X^\Delta_{\ell,t-7j},
\end{multline}
where we use \smash{$X^\Delta_{\ell,t} = (X_{\ell,t} - X_{\ell,
    t-7})/(X_{\ell,t-7})$}, and again $X_{\ell,t}$ stands for any of the five  
auxiliary indicators at location $\ell$ and time $t$.  We fit all models
\eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x} using logistic regression, 
pooling all 306 HRRs and using a 21-day trailing window for the training data.

@@

We begin with a summary of the high-level conclusions.  Here, and in the
following subsections, we will use ``AR'' to refer to the pure autoregressive
model both in forecasting \eqref{eq:forecast_ar} and in hotspot prediction
\eqref{eq:hotspot_ar} (where the reference to the task should be clear from the
context). We will also use the name of an auxiliary indicator---``CHNG-CLI'', 
``CHNG-CLI'', ``CHNG-COVID'', ``CTIS-CLI-in-community'', ``DV-CLI'', or
``Google-AA''---interchangeably with the model in forecasting
\eqref{eq:forecast_ar_x} or hotspot prediction \eqref{eq:hotspot_ar_x} that uses
this particular indicator as a feature (the meaning should be clear from the
context).  So, e.g., the CHNG-CLI model in forecasting is the one in
\eqref{eq:forecast_ar_x} that sets $X_{\ell,t}$ to be the value of the CHNG-CLI
indicator at location $\ell$ and time $t$, fits the coefficients based on the
training data, and makes its predictions by setting $X_{\ell,t}$ to the
appropriate CHNG-CLI values.  Finally, we use the term ``indicator model'' to 
broadly refer to any one of the ten models of the form \eqref{eq:forecast_ar_x},
\eqref{eq:hotspot_ar_x} (five in each of the forecasting and hotspot prediction
tasks).

\begin{itemize}
\item Stratifying predictions by the ahead value ($a=7,\ldots,21$), and
  aggregating results over the prediction period (early June through end of
  December 2020), each of the indicator models generally gives a boost in
  predictive accuracy over the AR model, in both the forecasting and hotspot 
  prediction tasks.  The gains in accuracy generally attenuate as the ahead 
  value grows.

@@

\item In a more detailed analysis of forecasting performance, the indicator
  models tend to be worse than AR when case rates are increasing (this is most
  notable in CHNG-CLI and DV-CLI), and better than AR when case rates are flat
  or decreasing (most notable in CHNG-COV and CTIS-CLI-in-community).
  More rarely does an indicator model tend to beat AR when case rates are
  increasing, but there appears to be some evidence of this for Google-AA. 
  % In an increasing period, when an indicator model is worse than AR, it tends
  % to underpredict relative to AR (its median forecast is lower).  In a flat
  % or decreasing  period, when an indicator model is better than AR, AR tends
  % to overpredict relative to it (the median AR forecast is higher).

@@

>>>>

We treat hotspot prediction as a binary classification problem and use a setup 
altogether quite similar to the forecasting setup described previously.
Separately for each $a=7,\ldots,28$, to predict $Z_{\ell,t+a}$, we consider a
simple logistic model:
\begin{equation}
\label{eq:hotspot_ar}
\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s}, s \leq t) \big) 
= \alpha^{a,\tau} + \sum_{j=0}^2 \^{a,\tau}_j Y^\Delta_{\ell,t-7j},
\end{equation}
where $\logit(p) = \log(p/(1-p))$, the log-odds of $p$.

In addition to this ``pure'' autoregressive model, we also consider five
logistic models of the form:  
\begin{multline}
\label{eq:hotspot_ar_x}
\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s}, X_{\ell,s}, s \leq t) \big)
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j} +  
\sum_{j=0}^2 \gamma^{a,\tau}_j X^\Delta_{\ell,t-7j},
\end{multline}
where we use \smash{$X^\Delta_{\ell,t} = (X_{\ell,t} - X_{\ell,
    t-7})/(X_{\ell,t-7})$}, and again $X_{\ell,t}$ stands for any of the five  
auxiliary indicators at location $\ell$ and time $t$.  We fit all models
\eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x} using logistic regression, 
pooling all 306 HRRs and using a 21-day trailing window for the training data.

@@

We begin with a summary of the high-level conclusions.  Here, and in the
following subsections, we will use ``AR'' to refer to the pure autoregressive
model both in forecasting \eqref{eq:forecast_ar} and in hotspot prediction
\eqref{eq:hotspot_ar} (where the reference to the task should be clear from the
context). We will also use the name of an auxiliary indicator---``CHNG-CLI'', 
``CHNG-CLI'', ``CHNG-COVID'', ``CTIS-CLI-in-community'', ``DV-CLI'', or
``Google-AA''---interchangeably with the model in forecasting
\eqref{eq:forecast_ar_x} or hotspot prediction \eqref{eq:hotspot_ar_x} that uses
this particular indicator as a feature (the meaning should be clear from the
context).  So, e.g., the CHNG-CLI model in forecasting is the one in
\eqref{eq:forecast_ar_x} that sets $X_{\ell,t}$ to be the value of the CHNG-CLI
indicator at location $\ell$ and time $t$, fits the coefficients based on the
training data, and makes its predictions by setting $X_{\ell,t}$ to the
appropriate CHNG-CLI values.  Finally, we use the term ``indicator model'' to 
refer to any one of the ten models of the form \eqref{eq:forecast_ar_x},
\eqref{eq:hotspot_ar_x} (five from each of the forecasting and hotspot
prediction tasks).

\begin{itemize}
\item Stratifying predictions by the ahead value ($a=7,\ldots,21$), and
  aggregating results over the prediction period (early June through end of
  December 2020), we find that each of the indicator models generally gives a
  boost in predictive accuracy over the AR model, in both the forecasting and
  hotspot prediction tasks.  The gains in accuracy generally attenuate as the
  ahead value grows.

@@

\item In a more detailed analysis of forecasting performance, we find that the
  indicator models tend to be worse than AR when case rates are increasing (this
  is most notable in CHNG-CLI and DV-CLI), and better than AR when case rates
  are flat or decreasing (most notable in CHNG-COVID and CTIS-CLI-in-community).
  More rarely does an indicator model tend to beat AR when case rates are 
  increasing, but there appears to be some evidence of this for the Google-AA
  model. 
  % In an increasing period, when an indicator model is worse than AR, it tends
  % to underpredict relative to AR (its median forecast is lower).  In a flat
  % or decreasing period, when an indicator model is better than AR, AR tends
  % to overpredict relative to it (the median AR forecast is higher).

@@

\subsection{Aggregate Results by Ahead Value}

\subsection{Implicit Regularization Hypothesis}

---which is that 
  defining features from lagged indicator values in the  models in
  \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar_x} simply provides a kind of
  implicit regularization on top of the basic AR model (so the benefits have
  nothing to do with the actual indicator values)---

To test the ``implicit regularization'' hypothesis (which we explain
  precisely later), we reran all of the prediction experiments but with
  $X_{\ell,t}$ in each indicator model replaced by suitable random noise.  This 
  resulted in vastly different (worse) performance compared to the indicator
  models fit with actual indicator values, and gave no improvement over the 
  pure AR model, in both forecasting and hotspot prediction, providing a strong
  rejection (informally speaking) of the implicit regularization hypothesis.

\subsection{Performance in Up, Down, Flat Periods}

\subsection{Leading-ness or Lagging-ness}

TODO

% TODO: make th

A Primary 
B Bootstrap
C up/down/flat (Dan's first histogram) --> table in appendix
D leading/lagging
